%!TEX root = ../main.tex
In this section, we define cost-enriched finite state automata (CEFA), which was introduced in \cite{atva2020} and will be used to solve the satisfiability problem of RECL later on. 
%
Intuitively, CEFA adds write-only cost registers to finite state automata. ``write-only'' means that the cost registers can only be written/updated but cannot be read, i.e., they cannot be used to guard the transitions. 
%
\begin{definition}[Cost-Enriched Finite Automaton]
  A cost-enriched finite automaton $\aut$ is a tuple $(R, Q, \Sigma, \delta, I, F, \alpha)$ where
  \begin{itemize}
  \item $R = \{r_1, \cdots, r_k\}$ is a finite set of registers, 
    \item $Q, \Sigma, I, F$ are as in the definition of NFA,
    \item $\delta \subseteq Q \times \Sigma \times Q \times \Int^R$ is a transition relation, where $\Int^R$ denotes the updates on the values of registers.
    \item $\alpha \in \Phi(R)$ is an LIA formula specifying an accepting condition.
  \end{itemize}
\end{definition}

For convenience, we use $R_\aut$ to denote the set of registers of $\aut$. We assume a linear order on $R$ and write $R$ as a vector $(r_1, \cdots, r_k)$. Accordingly, we write an element of $\Int^R$ as a vector $(v_1, \cdots, v_k)$, where $v_i$ is the update of $r_i$ for each $i \in [k]$. We also write a transition $(q, a, q', \vec{v}) \in \delta$ as $q \xrightarrow[\vec{v}]{a} q'$.

The semantics of CEFA is defined as follows. Let $\aut = (R, Q, \Sigma, \delta, I, F, \alpha)$ be a CEFA. 
A \emph{run} of $\aut$ on a string $w = a_1 \cdots a_n$ is a sequence $q_0 \xrightarrow[\myvec{v_1}]{a_1} q_1 \cdots q_{n-1}\xrightarrow[\myvec{v_n}]{a_n} q_n$ such that $q_0 \in I$ and $q_{i-1} \xrightarrow[\myvec{v_i}]{a_i} q_i$ for each $i \in [n]$. A run $q_0 \xrightarrow[\myvec{v_1}]{a_1} q_1 \cdots q_{n-1}\xrightarrow[\myvec{v_n}]{a_n} q_n$ is \emph{accepting} if $q_n \in F$ and $\alpha(\myvec{v'}/R)$ is true, where $\myvec{v'} = \sum \limits_{j \in [n]} \myvec{v_j}$. The string $w$ is $accepted$ by the CEFA if and only if there is an accepting run of $w$. The vector $\myvec{v'}= \sum \limits_{j \in [n]} \myvec{v_j}$ is called the \emph{cost} of an accepting run $q_0 \xrightarrow[\myvec{v_1}]{a_1} q_1 \cdots q_{n-1}\xrightarrow[\myvec{v_n}]{a_n} q_n$.Note that the values of all registers are initiated to zero and updated to $\sum \limits_{j \in [n]} \myvec{v_j}$ after all the transitions in the run are executed. We use $\myvec{v'} \in \aut(w)$ to denote the fact that there is an accepting run of $\aut$ on $w$ whose cost is $\myvec{v'}$.  
We define the semantics of a CEFA $\aut$, denoted by $\Lang(\aut)$, as $\{(w; \myvec{v'}) \mid  \myvec{v'} \in \aut(w)\}$.  In particular, if $I \cap F \neq \emptyset$ and $\alpha[\myvec{0}/R]$ is true, then $(\varepsilon; \myvec{0}) \in \Lang(\aut)$. Moreover, we define the \emph{output} of a CEFA $\aut$, denoted by $\cefaout(\aut)$, as $\{\myvec{v'} \mid  \exists w.\ \myvec{v'} \in \aut(w)\}$.

We want to remark that the definition of CEFA above is slightly different from that of \cite{atva2020}, where CEFA did not include accepting conditions.
Moreover, the accepting conditions $\alpha$ in CEFA are defined in a \emph{global} fashion because the accepting condition does not distinguish final states. This technical choice is made so that the determinization and minimization of NFA can be utilized to reduce the size of CEFA in the next section. 

In the sequel, we define three CEFA operations: union, intersection, and concatenation. The following section will use these three operations to solve RECL constraints. Note that the union, intersection, and concatenation operations are defined in a slightly more involved manner than register automata~\cite{ra} and counter automata~\cite{GGM12}, as a result of the (additional) accepting conditions. 
%the technical choice that the accepting condition is attached to each automaton, not to each final state.

Let $\aut_1 = (R_1, Q_1, \Sigma, \delta_1, q_{1,0}, F_1, \alpha_1)$ and  $\aut_2 = (R_2, Q_2, \Sigma, \delta_2, q_{2,0}, F_2, \alpha_2)$\footnote{\hdh{We use a single initial state for simplicity. A set of initial states can be transferred to a single one by introducing a new state $q_I$ and adding $\epsilon-$transition from $q_I$ to each initial state.}} be two CEFA that share the alphabet. 
Moreover, suppose that $R_1 \cap R_2 = \emptyset$ and $Q_1 \cap Q_2 = \emptyset$. 

The \emph{union} of $\aut_1$ and $\aut_2$ is denoted by $\aut_1 \cup \aut_2$. Two fresh auxiliary registers say $r'_1,r'_2 \not \in R_1 \cup R_2$, are introduced so that the accepting condition knows whether a run is from $\aut_1$ (or $\aut_2$).\hdh{This is mainly because the accepting condition is "global", i.e., the accepting condition is attached to the CEFA, not specific final state.} Specifically, $\aut_1 \cup \aut_2 = (R', Q', \Sigma, \delta', I', F', \alpha')$ where 
\begin{itemize}
\item $R' = R_1 \cup R_2 \cup \{r'_1, r'_2\}$, $Q' = Q_1 \cup Q_2 \cup \{q'_0\}$ with $q'_0 \not \in Q_1 \cup Q_2$, $I' = \{q'_0\}$, 
%
\item $\delta'$ is the union of four transitions sets:
\begin{itemize}
  \item $\{(q'_0, a, q'_1, (\myvec{v_1}, \myvec{0}, 1,0)) \mid \exists q_1 \in I_1.\ (q_1, a, q'_1, \myvec{v_1}) \in \delta_1\}$
  \item $\{(q'_0, a, q'_2, (\myvec{0}, \myvec{v_2}, 0, 1)) \mid \exists q_2 \in I_2.\ (q_2, a, q'_2, \myvec{v_2}) \in \delta_2\}$
  \item $\{(q_1, a, q'_1, (\myvec{v_1}, \myvec{0}, 0, 0)) \mid q_1\not\in I_1.\ (q_1, a, q'_1, \myvec{v_1}) \in \delta_1\}$
  \item $\{(q_2, a, q'_2, (\myvec{0}, \myvec{v_2}, 0, 0)) \mid q_2\not\in I_2.\ (q_2, a, q'_2, \myvec{v_2}) \in \delta_2\}$
\end{itemize}
 where $(\myvec{v_1}, \myvec{0}, 1,0)$ is a vector that updates $R_1$ by $\myvec{v_1}$, updates $R_2$ by $\myvec{0}$, and updates $r'_1,r'_2$ by $1,0$ respectively. Similarly for $(\myvec{0}, \myvec{v_2}, 0, 1)$, and so on.

\item $F'$ and $\alpha'$ are defined as follows, 
\begin{itemize}
\item if $(\epsilon; \myvec{0})$ belongs to $\lan(\aut_1)$ or $\lan(\aut_2)$, i.e., one of the two automata accepts the empty string $\epsilon$, then $F' = F_1 \cup F_2 \cup \{q'_0\}$ and $\alpha' = (r'_1 = 0 \wedge r'_2 = 0) \vee (r'_1=1 \wedge \alpha_1) \vee (r'_2=1 \wedge \alpha_2)$, 
\item otherwise, $F'=F_1 \cup F_2$ and $\alpha' = (r'_1=1 \wedge \alpha_1) \vee (r'_2=1 \wedge \alpha_2)$.
\end{itemize}
\end{itemize}
From the construction, we know that 
$$
\Lang(\aut_1 \cup \aut_2) = 
\left\{(w; \myvec{v}) \left \vert 
\begin{array}{l}
%\myvec{v} \in \Int^{R'}, 
(w; \prj_{R_1}(\myvec{v})) \in \Lang(\aut_1) \mbox{ and } \prj_{r_1'}(\myvec{v}) = 1, \mbox{ or } \\
(w; \prj_{R_2}(\myvec{v})) \in \Lang(\aut_2) \mbox{ and } \prj_{r_2'}(\myvec{v}) = 1, \mbox{ or } \\
(w; \myvec{v}) = (\epsilon, \myvec{0}) \mbox{ if } \aut_1 \mbox{ or } \aut_2 \mbox{ accepts } \epsilon
\end{array}
\right.
\right\}.
$$ 
Intuitively, $\aut_1 \cup \aut_2$ accepts the words that are accepted by one of the $\aut_1$ and $\aut_2$ and outputs the costs of the corresponding automaton.

The \emph{intersection} of $\aut_1$ and $\aut_2$, denoted by $\aut_1 \cap \aut_2 = (R', Q', \Sigma, \delta', I', F', \alpha')$, is defined in the sequel. 
\begin{itemize}
\item $R' = R_1 \cup R_2$, $Q' = Q_1 \times Q_2$, $I' = I_1 \times I_2$, $F' = F_1 \times F_2$, $\alpha' = \alpha_1 \wedge \alpha_2$, 
\item $\delta'$ comprises the tuples $((q_1, q_2), a, (q'_1, q'_2), (\myvec{v_1}, \myvec{v_2}))$ such that $(q_1,a, q'_1, \myvec{v_1}) \in \delta_1$ and $(q_2,a, q'_2, \myvec{v_2}) \in \delta_2$.
\end{itemize}
From the construction, 
$$\Lang(\aut_1 \cap \aut_2) = \{(w; \myvec{v}) \mid (w; \prj_{R_1}(\myvec{v})) \in \Lang(\aut_1) \mbox{ and } (w; \prj_{R_2}(\myvec{v})) \in \Lang(\aut_2)\}.$$
Intuitively, $\aut_1 \cap \aut_2$ accepts the words that are accepted by both $\aut_1$ and $\aut_2$ and outputs the costs of $\aut_1$ and $\aut_2$.  

The \emph{concatenation} of $\aut_1$ and $\aut_2$, denoted by $\aut_1 \concat \aut_2$, is defined similarly as that of NFA, that is, a tuple $(\hdh{R'}, Q', \Sigma, \delta', I', F', \alpha')$, where $\hdh{R' = R_1\cup R_2,} Q' = Q_1 \cup Q_2$, $I' = I_1$, $\alpha' = \alpha_1 \wedge \alpha_2$, $\delta' = \{(q_1, a, q'_1, (\myvec{v_1}, \myvec{0})) \mid (q_1, a, q'_1, \myvec{v_1}) \in \delta_1\} \cup \{(q_2, a, q'_2, (\myvec{0}, \myvec{v_2})) \mid (q_2, a, q'_2, \myvec{v_2}) \in \delta_2\} \cup \{(q_1, a, q_2, (\myvec{0}, \vec{v_2})) \mid q_1 \in F_1, \exists q' \in I_2.\ (q', a, q_2, \vec{v_2}) \in \delta_2\}$, moreover, if $I_2 \cap F_2 \neq \emptyset$, then $F'= F_1 \cup F_2$, otherwise, $F'= F_2$. From the construction, 
$$\Lang(\aut_1 \concat \aut_2) = \{(w_1w_2; \myvec{v}) \mid (w_1; \prj_{R_1}(\myvec{v})) \in \Lang(\aut_1) \mbox{ and }(w_2; \prj_{R_2}(\myvec{v})) \in \Lang(\aut_2)\}.$$

Furthermore, the union, intersection, and concatenation operations can be extended naturally to multiple CEFA, that is, $\aut_1 \cup \cdots \cup \aut_n$, $\aut_1 \cap \cdots \cap \aut_n$, $\aut_1 \concat \cdots \concat \aut_n$. For instance, $\aut_1 \cup \aut_2 \cup \aut_3 = (\aut_1 \cup \aut_2) \cup \aut_3$, $\aut_1 \cap \aut_2 \cap \aut_3 = (\aut_1 \cap \aut_2) \cap \aut_3$, and $\aut_1 \concat \aut_2 \concat \aut_3 = (\aut_1 \concat \aut_2) \concat \aut_3$.