%!TEX root = ../main.tex
The string data type plays a crucial role in modern programming languages such as JavaScript, Python, Java, and PHP. 
String manipulations are error-prone and could even give rise to severe security vulnerabilities (e.g., cross-site scripting, aka XSS). 
One powerful method for identifying such bugs is \emph{symbolic execution}, which is possibly in combination with dynamic analysis. It analyses symbolic paths in a program by viewing them as constraints checked by constraint solvers. 
%
Symbolic execution of string manipulating programs has motivated the highly active research area of \emph{string constraint solving}, resulting in the development of numerous string solvers in the last decade, e.g.,
Z3seq~\cite{z3seq}, CVC4/5~\cite{cvc4,cvc5}, Z3str/2/3/4~\cite{Z3-str,Z3-str2,Z3-str3,BerzishMurphy2021}, Z3str3RE~\cite{BD+23}, 
Z3-Trau~\cite{Z3-trau}\cite{z3trau}, OSTRICH~\cite{CHL+19}, Slent~\cite{WC+18}, among many others. 

Regular expressions (regex for short) and the string-length function are widely used in string-manipulating programs. \sout{According to the statistics from \cite{CS16,DCSL18,WS18}, regexes are used in about 30–40\% of Java, JavaScript, and Python software projects. Moreover, string-length occupies 78\% of the occurrences of string operations in 18 Javascript applications, according to the statistics from \cite{malware_detection_3_kudzu}.} \hdh{Specifically, regexes are used in about 30–40\% of Java, JavaScript, and Python software projects, while string-length occupies 78\% of the occurrences of string operations in 18 Javascript applications\footnote{According to the statistics from \cite{CS16,DCSL18,WS18,malware_detection_3_kudzu}.}. As a result, most of the aforementioned string constraint solvers support both regexes and string-length function, and specialized algorithms have been proposed to solve such string constraints efficiently (see e.g. \cite{LTR+15,BD+23}).} 

Counting (aka repetition) is a convenient feature in regexes that counts the number of matchings of sub-patterns. For instance, $a^{\{2, 4\}}$ specifies that $a$ occurs at least twice and at most four times, and $a^{\{2, \infty\}}$ specifies that $a$ occurs at least twice. 
Note that the Kleene star and the Kleene plus operator are special cases of counting, i.e., $a^*$ is equivalent to $a^{\{0,\infty\}}$ and $a^+$ is equivalent to $a^{\{1,\infty\}}$.
Counting is a frequently used feature of regexes \hdh{:}\sout{According to the statistics from \cite{CS16},} kleene star/plus occur in more than 70\% of 1,204 Python projects, while other forms of counting occur in more than 20\% of them\hdh{\footnote{According to the statistics from \cite{CS16}.}}. Therefore, an efficient analysis of string manipulating programs requires efficient solvers for string constraints containing regexes with counting\footnote{In the rest of this paper, for clarity, we use counting to denote expressions of the form $e^{\{m, n\}}$ and  $e^{\{m, \infty\}}$, but not $e^*$ or $e^+$.} and string-length function at least. \hdh{To the best of our knowledge, there is no such solver optimized for both counting and string-length.}

\sout{Nevertheless, }The aforementioned state-of-the-art string constraint solvers still suffer from such string constraints, especially when the counting and length bounds are large. For instance, none of the string solvers CVC5, Z3seq, Z3-Trau, Z3str3, Z3str3RE, and OSTRICH is capable of solving the following constraint within 120 seconds,
%, and Z3 can not solve it within 120 seconds.
%
\begin{equation}\label{eqn-running}
x \in (\Sigma \setminus a)^{\{1, 60\}} (\Sigma \setminus b)^{\{1, 60\}} (\Sigma \setminus c)^{\{0, 60\}} \wedge x \in \Sigma^* c^+ \wedge |x| > 120.
\end{equation}
Intuitively, the constraint in (\ref{eqn-running}) specifies that $x$ is a concatenation of three strings $x_1, x_2, x_3$ where $a$ (resp. $b$, $c$) does not occur in $x_1$ (resp. $x_2$, $x_3$), moreover, $x$ ends with a nonempty sequence of $c$'s, and the length of $x$ is greater than $120$. It is easy to observe that this constraint is unsatisfiable since on the one hand, $|x| > 120$ and the counting upper bound $60$ in both $(\Sigma \setminus a)^{\{1, 60\}}$ and $(\Sigma \setminus b)^{\{1, 60\}}$ imply that $x$ must end with some character from $\Sigma \setminus c$, that is, a character different from $c$, and on the other hand, $x \in \Sigma^*c^+$ requires that $x$ has to end with $c$.

A typical way for string constraint solvers to deal with regular expressions with counting is to unfold them into those \emph{without} counting using the concatenation operator. For instance, $a^{\{1, 4\}}$ is unfolded into $a(\varepsilon + a + aa + aaa)$ and $a^{\{2,\infty\}}$ is unfolded into $aaa^{*}$. Since the unfolding incurs an exponential blow-up on the sizes of constraints (assuming that the counting in string constraints are encoded in binary), the unfolding penalizes the performance of the solvers, especially when the length bounds are also available.

\medskip
\noindent 
\emph{Contribution.} In this work, we focus on the class of string constraints with regular membership and string-length function, where the counting operators may occur (called RECL for brevity). We make the following contributions.
\begin{itemize}
  \item We propose an automata-theoretical approach for solving RECL constraints.
 Our main idea is to represent the counting operators by cost registers in cost-enriched finite automata (CEFA, see Section~\ref{sec:automaton} for the definition), instead of unfolding them explicitly. The string-length function is modeled by cost registers as well. The satisfiability of RECL constraints is reduced to the non-emptiness problem of CEFA w.r.t. a linear integer arithmetic (LIA) formula. According to the results from~\cite{atva2020}, an LIA formula can be computed to represent the potential values of registers in CEFA.
Thus, the non-emptiness of CEFA w.r.t. LIA formulas can be reduced to the satisfiability of LIA formulas, which is then tackled by off-the-shelf SMT solvers. Moreover, we propose techniques to reduce the sizes (i.e., the number of states and transitions) of CEFA in order to achieve better performance.
%
  \item For complex RECL constraints where the counting operators may be nested with the other counting operators or complement operators, we propose optimization techniques to improve the performance. Furthermore, for the satisfiable RECL constraints where the solutions of the LIA formulas can be obtained, but the models of the original RECL constraint cannot be generated,  we propose optimization techniques to accelerate the model generation process. 
%
%  \item 
%  Combined with the size-reduction techniques mentioned above, the register representation of regex-counting and string-length in CEFA entails an efficient algorithm for solving RECL constraints. 
 %
\item We implement the algorithm on top of OSTRICH, resulting in a string solver called $\ostrichrecl$. Finally, we utilize a collection of benchmark suites comprising 49,843 problem instances in total to evaluate the performance of $\ostrichrecl$. The experiment results show that $\ostrichrecl$ solves the RECL constraints more efficiently than the state-of-the-art string solvers (see Table~\ref{tab:table_overall_eval}-\ref{tab:results_model-gen}). For instance, $\ostrichrecl$ solves 49,638 out of 49,843 instances, while the best state-of-the-art solver solve only 47,721 instances. We also do experiments to justify some of our technical choices and evaluate the performance of the optimizations\footnote{\hdh{The tool is available as open source at \url{https://github.com/SimpleXiaohu/ostrich/tree/jsa2024}.

The benchmark suites are available at \url{https://github.com/SimpleXiaohu/ostrich/tree/jsa2024/zaligvinder/models/jsa2024}.}}. 
%
% 1,000 benchmark instances where the counting and length bounds are large or the counting operators are nested with some other counting operators or complement operators, $\ostrichrecl$ solves at least $278$ more instances than the other solvers, while spending only half or less time per instance on average. 
\end{itemize}

\medskip
\noindent
\emph{Related work.} 
We discuss more related work on regexes with counting, string-length function, and automata with registers/counters.  
%
Determinism of regexes with counting was investigated in~\cite{GGM12,CL15}. 
%
Real-world regexes in programming languages include features beyond classical regexes, e.g., the greedy/lazy Kleene star, capturing groups, and back references. Real-world regexes have been addressed in symbolic execution of Javascript programs~\cite{LMK19} and string constraint solving~\cite{CF+22}. 
Nevertheless, the counting operators are still unfolded explicitly in~\cite{CF+22}. 
%
The Trau tool focuses on string constraints involving flat regular languages and string-length function and solves them by computing LIA formulas that define the Parikh images of flat regular languages~\cite{z3trau}. The Slent tool solves the string constraints involving string-length function by encoding them into so-called length-encoded automata, then utilizing symbolic model checkers to solve their non-emptiness problem~\cite{WC+18}. However, neither Trau nor Slent supports counting operators explicitly, in other words, counting operators in regexes should be unfolded before solved by them.
%
Cost registers in CEFAs are different from registers in (symbolic) register automata~\cite{ra,sra}:  In register automata, registers are used to store input values and can only be compared for equality/inequality, while in CEFAs, cost registers are used to store integer-valued costs and can be updated by adding/subtracting integer constants and constrained by the accepting conditions which are LIA formulas.
%
Therefore, cost registers in CEFAs are more like counters in counter automata/machines~\cite{Minsky67}, that is, CEFAs can be obtained from counter machines by removing transition guards and adding accepting conditions. \hdh{CEFAs can also be considered as a variation of Parikh automata \cite{parikh_automaton_2003} in which the numbers in the alphabet can be integers, instead of natural numbers.}
%
%For the automata model, the definition of registers of CEFAs resembles the counters of the counter automata class~\cite{Minsky67}, but CEFAs contain no guards in each transition. 
%
Counting-set automata were proposed in~\cite{redos_lenka,HS+23} to quickly match a subclass of regexes with counting. Moreover, a variant of nondeterministic counter automata, called bit vector automata, was proposed recently in \cite{GKM23} to enable fast matching of a more expressive class of regexes with counting.   Nevertheless, the non-emptiness problem of these automata was not considered, and it is unclear whether these automata models can be used for solving string constraints with regex-counting and string-length. 

%Except for the string solvers mentioned before, some related works about string constraints with the length function exist. 

\smallskip
\noindent
\emph{Structure.} 
The rest of this paper is structured as follows: \sout{Section~\ref{sec:overview} gives an overview of the approach in this paper. Section~\ref{sec:pre} introduces the preliminaries.} \hdh{Section~\ref{sec:pre} introduces the preliminaries. Section~\ref{sec:overview} gives an overview of the approach in this paper.}
Section~\ref{sec:recl} presents the syntax and semantics of RECL. 
Section~\ref{sec:automaton} defines CEFA. Section~\ref{sec:algorithm} introduces the algorithm to solve RECL constraints. 
Section~\ref{sec:heuristic} presents some heuristics for improving the performance of solving RECL constraints where the counting operators are nested with some other counting operators or complement operators. Section~\ref{sec:implementation} shows the experiment results. Section~\ref{sec:conclu} concludes this paper.

\smallskip

A preliminary version of this paper has appeared in the proceedings of SETTA 2023~\cite{Denghang2023}. This paper extends the SETTA paper in the following three aspects.
\begin{itemize}
\item To improve the performance of complex RECL constraints where the counting operators are nested with some other counting operators or complement operators, various optimizations are proposed. 
\item To deal with the issue that the satisfiability of LIA formulas can be solved, but models of the original RECL constraints cannot be generated, we propose techniques to optimize the model generation process.  
\item The benchmarks are extended to cover the situations where the counting operators are nested with some other counting operators or complement operators, and the experiments are extended accordingly.  
\end{itemize}